\documentclass{homework}
\input{preamble}
\usepackage{hyperref}
\author{Jim Fowler}
\course{Math 5520H}
\title{Distributions}
\date{November 19, 2019}
\newcommand{\C}{\mathbb{C}}
\newcommand{\D}{\mathcal{D}}
\DeclareMathOperator{\image}{im}
\begin{document}
\maketitle

\begin{problem}
  We say $\varphi : \R \to \R$ is a  \textbf{test function} if $\varphi$ is smooth and has \textbf{compact support}.

  Does the set of test functions $\D(\R)$ form a vector space?
\end{problem}

\vfill

\begin{problem}
  A \textbf{distribution} is a linear functional $F : \D(\R) \to \R$ which
  is continuous in the sense that
  for every
  compact $K \subset \mathbb{R}$ there are constants $M$ and $m$ so
  that
  \[
    F(\varphi) \leq M \cdot \max_{n \leq m} \max_{x \in \R} | f^{(n)}(x) |.
  \]
  for $\varphi$ with support in $K$.

  %, in the sense that $\lim_{k\to\infty} F(\varphi_k) = F(\lim_{k\to\infty} \varphi_k)$.

  Verify that the \textbf{Dirac delta} defined by
  $\langle \delta, \varphi \rangle = \varphi(0)$ is a distribution.
\end{problem}

\vfill

\begin{problem}
  Suppose $f : \R \to \R$ is a locally integrable function.  (Define
  \textbf{locally integrable}.)

  Does $\langle f, \varphi \rangle = \int_\R f(x) \, \varphi(x) \, dx$
  define a distribution?  %We therefore identify certain functions with a corresponding distribution.
\end{problem}

\vfill

\begin{problem}
  Use integration by parts to define the derivative of a distribution.
\end{problem}

\vfill

\begin{problem}
  Define the \textbf{Heaviside step function} by
  \[
    H(x) = \begin{cases}
      0 &\mbox{if $x < 0$}, \\
      1 &\mbox{if $x \geq 0$}.
    \end{cases}
  \]
  Compute $H'(x)$.
\end{problem}

\vfill

\begin{problem}
  Let $f(x) = |x|$ and compute $f'(x)$ as a distribution.
\end{problem}

\vfill

\begin{problem}
  Compute the derivative of the Dirac delta function, i.e.,
  compute $\langle \delta', \varphi \rangle$.
\end{problem}

\vfill

\begin{problem}
  Does the distribution $\delta'$ arise from integrating against some function?
\end{problem}

\vfill

\begin{problem}
  For a distribution $f$ and a smooth function $\psi$, define a distribution $f \psi$.
  %by the rule $\langle f \psi, \varphi \rangle = \langle f, \psi \varphi$.  Does this definition make sense?
\end{problem}

\vfill

\begin{problem}
  Compute $\delta \cdot x$.
\end{problem}

\vfill

\begin{problem}
  Define $f(x) = x (\log   |x|  — 1)$ and verify $f'(x) = \log |x|$.
\end{problem}

\vfill

% http://sites.mathdoc.fr/OCLS/pdf/OCLS_1954__21__1_0.pdf

\begin{problem}
  Verify further that $f''(x) \cdot x = 1$.
\end{problem}

\vfill

% f'(x) = (\log   |x|  — 1) + x (1/x)

\begin{problem}
  Can you multiply distributions in general?  Consider
  \[
    0 = x^{-1} \cdot 0 = x^{-1} (x \delta) = (x^{-1} x) \delta = 1 \cdot \delta = \delta.
  \]
  This is the \textbf{Schwartz impossibility theorem.}
\end{problem}

\vfill

%https://math.stackexchange.com/questions/1357877/schwartz-impossibility-result

  \end{document}
%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
