\documentclass{homework}
\input{preamble}
\usepackage{hyperref}
\author{Jim Fowler}
\course{Math 5520H}
\title{Assigned Readings for Week 12}
\date{Monday, November 5, 2018}
\begin{document}
\maketitle

At the end of this week, we will have \textbf{Midterm~2}, and since
you will also be studying for this challenging exam this week, I want
to make sure that you have time to reflect on your past learning and
that you have time to make connections between the linear algebra and
differential equations content.

The exam is built again from ten questions, consisting of one
definition, four numerical computations, three exploration problems,
and two ``prove or disprove, and salvage if possible'' prompts.
\textbf{Be sure to provide a response to every question asked} because
the only way to earn points is to provide responses.

This exam is cumulative, but of course overwhelmingly focuses on the
content covered since Midterm~1.  \textbf{In particular, new material
  from this week will not appear on Midterm~2}, so the upcoming exam
will focus on material from the ``eigenstuff'' in Week~5 all the way
to the techniques for solving differential equations with regular
singular points.

Our time in lecture will focus on digging into some linear algebra
which we have seen shadows of, but not yet discussed in depth in
class.  One of these topics are various constructions for new vector
spaces; we've seen the direct sum $V \oplus W$, and we've seen how
$\mathrm{Hom}(V,W)$ is a vector space, but perhaps you have not
yet thought carefully about the difference between
\[
  \prod_{k=1}^\infty V_k \mbox{ and }   \bigoplus_{k=1}^\infty V_k.
\]
Another construction is that of \textbf{quotient spaces}, which you
can review in Axler's \textit{Linear Algebra Done Right} in
\textsection 3.E Products and Quotients of Vector Spaces, or in
Curtis' \textit{Linear Algebra: An Introductory Approach} in
\textsection 26 Quotient spaces and dual vector spaces.

\textbf{Duality} is another topic that provides opportunities to
reinforce your study of linear algebra.  From \textit{Linear Algebra
  Done Right}, study \textsection 3.F Duality; from \textit{Linear
  Algebra: An Introductory Approach}, study the aforementioned
\textsection 26 and also \textsection 27 Bilinear forms and duality.
For our ``computational'' task this week, I would like you to start
with some reasonable basis (say, a basis of monomials for polynomials
of bounded degree) and find a dual basis.

Through duality, we will be led to thinking more carefully about
\textbf{indices} which will lead us to brief foray into tensor
products $V \otimes W$ and more \textbf{multilinear algebra}, making
connections back to the \textbf{wedge products} we first saw when
thinking about a basis-free description of the \textbf{determinant.}

At the end of the week, with duality, we will study linear functionals
on spaces of polynomials, i.e., \textbf{the umbral calculus}, so-named
for its power in proving ``shadowy'' identities involving polynomials, e.g.,
behold the syntactic similarity between
\[
 \frac {d}{dx} x^{n}=n\,x^{n-1} \mbox{ and } {\frac {d}{dx}}B_{n}(x)=n \, B_{n-1}(x),
\]
where $B_n$ denotes the $n^{\mbox{\scriptsize th}}$ Bernoulli polynomial.

\end{document}
